---
title: "math_340_final_project_NFTs"
author: "Robert Alward"
date: "11/20/2021"
output: html_document
---

```{r setup, include=FALSE}
# Libraries
library(tidyverse)
#library(anytime)
library(foreign)
library(ggplot2)
library(MASS)
library(car)
library(leaps)
library(stats)
library(ISLR)
library(glmnet)
library(randomForest)
library(caTools)
library(varImp)
```

*General Data*
```{r}
combo_df <- read.csv("combo_df.csv")
valid_df <- read.csv("valid_combo_df.csv")
no_outlier_df <- combo_df[-c(21, 32, 39), ] 
```

*Test Data*
```{r}
test_x <- valid_df[,c(6,8,9,10,11,12,13,14,15)]
test_y <- valid_df[,c(16)]
```

*Log Transformed Data*
```{r}
combo_df_log <- combo_df[,c(6,8,9,10,11,12,13,14,15,16)] + 1
combo_df_log <- log(combo_df_log[,2:10])
combo_df_log$diff <- combo_df$diff
```

*Test Data*
```{r}
test_x <- valid_df[,c(6,8,9,10,11,12,13,14,15)]
test_y <- valid_df[,c(16)]
```

*Log Transformed Data*
```{r}
test_df_log <- valid_df[,c(6,8,9,10,11,12,13,14,15,16)] + 1
test_df_log <- log(test_df_log[,2:10])
test_df_log$diff <- valid_df$diff
test_x_log <- test_df_log[,c(1:8,10)]
test_y_log <- test_df_log[,9]
```
### *Model Initialization*

# 1. Multiple Linear Regression
```{r}
full.lm <- lm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data = combo_df)
summary(full.lm)
```

# 2 Robust Linear Regression
```{r}
full.rlm <- rlm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data = combo_df)
summary(full.rlm)
```

# 3 Ridge Regression
```{r}
full.ridge <- lm.ridge(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data=combo_df, lambda=seq(0, 30, 1e-3))

best.ridge <- lm.ridge(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data=combo_df, lambda=.11)

#summary(full.ridge)
select(full.ridge)
names(best.ridge)
best.ridge$coef
```

# 4 Weighted Least Squares Regression
```{r}
#Data prep for WLS regression
sqr_resid <- full.lm$residuals^2
combo_df_wls <- cbind(combo_df,sqr_resid)

wls_practice.lm <- lm(sqr_resid ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data=combo_df_wls)

full.wls <- lm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data=combo_df_wls, weights=(1/(wls_practice.lm$fit)^2))
summary(full.wls)
```

# 5 Modeling with Random forest 
```{r}
combo_df_t = combo_df[,c(6,8,9,10,11,12,13,14,15,16)]
rf <- randomForest(
  current_mean_val ~ .,
  data=combo_df_t,
  importance = T
)
```


# 6. Three Predictor Linear Model
```{r}
three.lm <- lm(current_mean_val ~ diff + mean_trans_val + mint_mean_val, data = combo_df)
summary(three.lm)
```

# 7. Eight Predictor Linear Model
```{r}
eight.lm <- lm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_mean_val, data = combo_df)
summary(eight.lm)
```

# 8. Linear Model: Mean Transfer Value Interaction
```{r}
interaction.trans.lm <- lm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val + mean_trans_val*diff + mean_trans_val*count_transfers + mean_trans_val*count_unique_trans + mean_trans_val*count_unique_owners + mean_trans_val*count_nfts + mean_trans_val*mint_mean_val + mean_trans_val*mint_min_val+ mean_trans_val*mint_max_val, data = combo_df)
summary(interaction.trans.lm)
```

# 9. Linear Model: Mean Mint Value Interaction
```{r}
interaction.mint.lm <- lm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val + mint_mean_val*diff + mint_mean_val*count_transfers + mint_mean_val*count_nfts , data = combo_df)
summary(interaction.mint.lm)
```

# 10. Log Linear Model
```{r}
log.lm <- lm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val , data = combo_df_log)
summary(log.lm)
```

# 11. Transfer Linear Model
```{r}
transfer.lm <- lm(current_mean_val ~ mean_trans_val , data = combo_df)
summary(transfer.lm)
```

# 12. No Transfer Linear Model
```{r}
no.transfer.lm <- lm(current_mean_val ~ diff + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data = combo_df)
summary(no.transfer.lm)
```



#### *Mean Squared Error*

# Evaluation Data
```{r}
valid_test <- valid_df[,c(6,8,9,10,11,12,13,14,15)]
valid_y <- valid_df[,c(16)]
```

# 1. Linear Model: Mean Squared Error Evaluation
```{r}
pred_lm <- predict(full.lm, newdata = valid_test)
lm_eval <- data.frame(pred_lm, actual = valid_y)
full.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("Linear MSE")
full.lm.mse/100000000
print("RMSE")
sqrt(full.lm.mse)
print("AIC")
AIC(full.lm)
print("BIC")
BIC(full.lm)
print("Adj. R-squared")
summary(full.lm)$adj.r.squared
```

# 2. Robust Linear Regression: Mean Squared Error Evaluation
```{r}
pred_rlm = predict(full.rlm, newdata = valid_test )
rlm_eval <- data.frame(pred_rlm, actual = valid_y)
full.rlm.mse <- mean((rlm_eval$pred_rlm - rlm_eval$actual)^2)
print("RLM MSE")
full.rlm.mse/100000000
print("RMSE")
sqrt(full.rlm.mse)
print("AIC")
AIC(full.rlm)
print("BIC")
BIC(full.rlm)
print("Adj. R-squared")
summary(full.rlm)$adj.r.squared
```

# 3. Ridge Regression: Mean Squared Error Evaluation
```{r}
predicted = full.ridge$ym +
  scale(test_x, center = full.ridge$xm, scale = full.ridge$scales) %*%
  full.ridge$coef[, which.min(full.ridge$GCV)]
as.numeric(predicted)
```

```{r}
ridge.mse <- mean((predicted - test_y)^2)
print("Ridge MSE")
ridge.mse/100000000
print("RMSE")
sqrt(ridge.mse)
print("AIC")
AIC(ridge.lm)
print("BIC")
BIC(ridge.lm)
print("Adj. R-squared")
summary(ridge.lm)$adj.r.squared
```

# 4. Weighted Least Squares Model: Mean Squared Error Evaluation
```{r}
pred_wls = predict(full.wls, newdata = valid_test)
wls_eval <- data.frame(pred_wls, actual = valid_y)
full.wls.mse <- mean((wls_eval$pred_wls - wls_eval$actual)^2)
print("WLS MSE")
full.wls.mse/100000000
print("RMSE")
sqrt(full.wls.mse)
print("AIC")
AIC(full.wls)
print("BIC")
BIC(full.wls)
print("Adj. R-squared")
summary(full.wls)$adj.r.squared
```

# 5. Random Forest Model: Mean Squared Error Evaluation
```{r}
pred_rf = predict(rf, newdata = valid_test )
rf_eval <- data.frame(pred_rf, actual = valid_y)
full.rf.mse <- mean((rf_eval$pred_rf - rf_eval$actual)^2)
print("RF MSE")
full.rf.mse/100000000
print("RMSE")
sqrt(full.rf.mse)
```

# 6. Three Predictor Linear Model: Mean Squared Error Evaluation
```{r}
pred_lm <- predict(three.lm, newdata = valid_test)
lm_eval <- data.frame(pred_lm, actual = valid_y)
three.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("Three Linear MSE")
three.lm.mse/100000000
print("RMSE")
sqrt(three.lm.mse)
print("AIC")
AIC(three.lm)
print("BIC")
BIC(three.lm)
print("Adj. R-squared")
summary(three.lm)$adj.r.squared
```

# 7. Eight Predictor Linear Model: Mean Squared Error Evaluation
```{r}
pred_lm <- predict(eight.lm, newdata = valid_test)
lm_eval <- data.frame(pred_lm, actual = valid_y)
eight.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("Eight Linear MSE")
eight.lm.mse/100000000
print("RMSE")
sqrt(eight.lm.mse)
print("AIC")
AIC(eight.lm)
print("BIC")
BIC(eight.lm)
print("Adj. R-squared")
summary(eight.lm)$adj.r.squared
```

# 8. Transfer Interaction Linear Model
```{r}
pred_lm <- predict(interaction.trans.lm, newdata = valid_test)
lm_eval <- data.frame(pred_lm, actual = valid_y)
transferint.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("Transfer Interaction Linear MSE")
transferint.lm.mse/100000000
print("RMSE")
sqrt(transferint.lm.mse)
print("AIC")
AIC(interaction.trans.lm)
print("BIC")
BIC(interaction.trans.lm)
print("Adj. R-squared")
summary(interaction.trans.lm)$adj.r.squared
```

# 9. Mint Interaction Linear Model: Mean Squared Error Evaluation
```{r}
pred_lm <- predict(interaction.mint.lm, newdata = valid_test)
lm_eval <- data.frame(pred_lm, actual = valid_y)
mintint.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("Mint Interaction Linear MSE")
mintint.lm.mse/100000000
sqrt(mintint.lm.mse)
print("RMSE")
sqrt(mintint.lm.mse)
print("AIC")
AIC(interaction.mint.lm)
print("BIC")
BIC(interaction.mint.lm)
print("Adj. R-squared")
summary(interaction.mint.lm)$adj.r.squared
```

# 10. Log Linear Model: Mean Squared Error Evaluation

```{r}
log_pred_lm <- predict(log.lm, newdata = test_x_log)
pred_lm <- exp(log_pred_lm)

lm_eval <- data.frame(pred_lm, actual = valid_y)
log.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("Log Linear MSE")
log.lm.mse/100000000
print("RMSE")
sqrt(log.lm.mse)
print("AIC")
AIC(log.lm)
print("BIC")
BIC(log.lm)
print("Adj. R-squared")
summary(log.lm)$adj.r.squared
```

# 11. Only Transfer Linear Model: Mean Squared Error Evaluation
```{r}
pred_lm <- predict(transfer.lm, newdata = valid_test)
lm_eval <- data.frame(pred_lm, actual = valid_y)
transfer.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("Only Transfer Linear MSE")
transfer.lm.mse/100000000
print("RMSE")
sqrt(transfer.lm.mse)
print("AIC")
AIC(transfer.lm)
print("BIC")
BIC(transfer.lm)
print("Adj. R-squared")
summary(transfer.lm)$adj.r.squared
```

# 12. No Transfer Linear Model: Mean Squared Error Evaluation
```{r}
pred_lm <- predict(no.transfer.lm, newdata = valid_test)
lm_eval <- data.frame(pred_lm, actual = valid_y)
no.transfer.lm.mse <- mean((lm_eval$pred_lm - lm_eval$actual)^2)
print("No Transfer Linear MSE")
no.transfer.lm.mse/100000000
print("RMSE")
sqrt(no.transfer.lm.mse)
print("AIC")
AIC(no.transfer.lm)
print("BIC")
BIC(no.transfer.lm)
print("Adj. R-squared")
summary(no.transfer.lm)$adj.r.squared
```