---
title: "Math 340 Final Project EDA Outputs"
author: "Robert Alward"
date: "12/09/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(foreign)
library(ggplot2)
library(MASS)
library(car)
library(leaps)
library(stats)
library(ISLR)
library(glmnet)
library(randomForest)
library(caTools)
library(varImp)
```

### 1.a Data loading

```{r}
combo_df <- read.csv("combo_df.csv")
valid_df <- read.csv("valid_combo_df.csv")
no_outlier_df <- combo_df[-c(21, 32, 39), ] 
```

```{r}
combo_df_log <- combo_df[,c(6,8,9,10,11,12,13,14,15,16)] + 1
combo_df_log <- log(combo_df_log[,2:10])
```


### 1.b Exploratory Data Analysis

*Correlation Matrix*
```{r}
cor_chart <- cor(combo_df[,c(2,4,6,8,9,10,11,12,13,14,15,16)])
cor_chart
str(as.data.frame(cor_chart))
write.csv(as.data.frame(cor_chart),"correlation_chart.csv")
```

*Correlation Plot*
```{r}
pairs(combo_df[,c(6,8,9,10,11,12,13,14,15,16)], pch = 19)
```

*Boxplots*
```{r}
par(mfrow = c(2,2))
boxplot(combo_df$diff_hr, main="Time Difference from First Minitng \n to Most Recent Transfer", ylab="Time")
boxplot(combo_df$mean_trans_val, main="Mean Value During NFT Transfer", ylab="microether (10^-6 ETH)")
boxplot(combo_df$count_transfers,main="Count of NFT Transfers", ylab="count")
boxplot(combo_df$count_unique_trans,main="Count of Transfers of Unique NFTs", ylab="count")
boxplot(combo_df$count_unique_owners,main="Count of Unique Owners of NFTs", ylab="count")
boxplot(combo_df$count_nfts,main="Count of Copies of an NFT", ylab="count")
boxplot(combo_df$mint_min_val,main="Minimum Value of NFT Copy at Miniting", ylab="microether (10^-6 ETH)")
boxplot(combo_df$mint_max_val,main="Maximum Value of NFT Copy at Miniting", ylab="microether (10^-6 ETH)")
boxplot(combo_df$mint_mean_val,main="Mean Value of NFT at Miniting", ylab="microether (10^-6 ETH)")
boxplot(combo_df$current_mean_val,main="Current Mean Value of NFT", ylab="microether (10^-6 ETH)")
```

*Outlier Dataset based on Mean Transfer Value*
```{r}
# Mean transfer value outliers
out <- boxplot.stats(combo_df$mean_trans_val)$out
out_ind <- which(combo_df$mean_trans_val %in% c(out))
out_df <- combo_df[out_ind,]
write.csv(as.data.frame(out_df),"outlier_df.csv")
```

*Histogram*
```{r}
par(mfrow = c(2,2))
hist(combo_df$diff_hr,main="Time Difference from First Minting \n to Most Recent Transfer", xlab="Time(seconds)", breaks = 20)
hist(combo_df$mean_trans_val,main="Mean Value During NFT Transfer", xlab="microether (10^-6 ETH)", breaks = 20)
hist(combo_df$count_transfers,main="Count of NFT Transfers", xlab="count", breaks = 20)
hist(combo_df$count_unique_trans,main="Count of Transfers of Unique NFTs", xlab="count",breaks = 20)
hist(combo_df$count_unique_owners,main="Count of Unique Owners of NFTs",xlab="count", breaks = 20)
hist(combo_df$count_nfts,main="Count of Copies of an NFT", xlab="count",breaks = 20)
hist(combo_df$mint_min_val,main="Minimum Value of NFT Copy at Miniting",xlab="microether (10^-6 ETH)", breaks = 20)
hist(combo_df$mint_max_val,main="Maximum Value of NFT Copy at Miniting",xlab="microether (10^-6 ETH)",breaks = 20)
hist(combo_df$mint_mean_val,main="Mean Value of NFT at Miniting", xlab="microether (10^-6 ETH)",breaks = 20)
hist(combo_df$current_mean_val,main="Current Mean Value of NFT", xlab="microether (10^-6 ETH)",breaks = 20)
```
*Summary Statistics*

```{r}
summary(combo_df$diff_hr,breaks = 20)
summary(combo_df$mean_trans_val,breaks = 20)
summary(combo_df$count_transfers,breaks = 20)
summary(combo_df$count_unique_trans,breaks = 20)
summary(combo_df$count_unique_owners,breaks = 20)
summary(combo_df$count_nfts,breaks = 20)
summary(combo_df$mint_min_val,breaks = 20)
summary(combo_df$mint_max_val,breaks = 20)
summary(combo_df$mint_mean_val,breaks = 20)
summary(combo_df$current_mean_val,breaks = 20)
```

*Residual Plots*

```{r}
# Multiple Linear Regression
full.lm <- lm(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data = combo_df)
plot(full.lm)
```

*Shapiro Test*
```{r}
shapiro.test(full.lm$residuals)
```


*Variance Inflation Factor*
```{r}
full.lm.vif <- vif(full.lm)
full.lm.vif
mean(full.lm.vif)
max(full.lm.vif)
```
The Max VIF is higher than desired as it is over 10
The Mean VIF is also very high (18.305) indicating a robust linear regression could be a good approach

*Ridge Trace Plots*
```{r}
big_lam <- seq(0,1,.01)
full.ridge <- lm.ridge(current_mean_val ~ diff + mean_trans_val + count_transfers + count_unique_trans + count_unique_owners + count_nfts + mint_min_val + mint_max_val + mint_mean_val, data=combo_df, lambda=big_lam)
select(full.ridge)

plot(full.ridge)
```
*Plot of Current Mean Value and Squared Residuals*

```{r}
plot(combo_df$current_mean_val,full.lm$residuals^2, main = "Plot of Current Mean Value and Squared Residuals",xlab = "Current Mean Value (microether (10^-6 ETH))", ylab = "Squared Residuals")
```


*Log Transformed Data*
```{r}
hist(combo_df_log$mean_trans_val,main="Mean Value During NFT Transfer", xlab="microether (10^-6 ETH)", breaks = 20)
hist(combo_df_log$count_transfers,main="Count of NFT Transfers", xlab="count", breaks = 20)
hist(combo_df_log$count_unique_trans,main="Count of Transfers of Unique NFTs", xlab="count",breaks = 20)
hist(combo_df_log$count_unique_owners,main="Count of Unique Owners of NFTs",xlab="count", breaks = 20)
hist(combo_df_log$count_nfts,main="Count of Copies of an NFT", xlab="count",breaks = 20)
hist(combo_df_log$mint_min_val,main="Minimum Value of NFT Copy at Miniting",xlab="microether (10^-6 ETH)", breaks = 20)
hist(combo_df_log$mint_max_val,main="Maximum Value of NFT Copy at Miniting",xlab="microether (10^-6 ETH)",breaks = 20)
hist(combo_df_log$mint_mean_val,main="Mean Value of NFT at Miniting", xlab="microether (10^-6 ETH)",breaks = 20)
hist(combo_df_log$current_mean_val,main="Log Transformed Current Mean Value of NFT", xlab="microether (10^-6 ETH)",breaks = 20)
```


